A distributed key-value store using Raft

### 2014-06-05 proj/paper.tex
paper done

This script demonstrates the partition tolerance of the data store. More specifically, it retains consistency while experiencing a network partition, while suffering brief loss of availability during the leader changeover (with potentially unbounded duration of the changeover process).
\section{Issues and lessons learned}
\end{document}

>>>>

This script demonstrates the partition tolerance of the data store. More specifically, it retains consistency while experiencing a network partition, while suffering brief loss of availability during the leader changeover (with potentially unbounded duration of the changeover process).
\section{Issues and lessons learned}
\subsection{Implementation}
The biggest issue that I had here was figuring out exactly how the event-driven model of the Tornado IOLoop works and how to integrate timeouts with it. Most networked programs I've written before were threaded and operated on the socket level, which means that I had control over how timeouts and message-driven events interact. The main concern I had was race conditions caused by timeouts triggering in the middle of a function execution. Fortunately, as a read over the Tornado source code revealed, the triggered timeouts were simply placed on a queue for later execution, so it turned out to be a non-issue.

In addition, the design of the broker did not allow for very effective debugging. Since all the output got redirected to \verb|/dev/null|, I was unable to track down Python exceptions that got thrown when I messed up with indexing. I eventually resolved this problem by telling the interpreter to use a log file in \verb|/tmp| as stdout and stderr in the code for initializing the nodes. Beyond these two issues, the implementation process was fairly straightforward. Python and ZMQ make it very easy to abstract away the gritty details of a network, and that reduced the workload considerably compared to if I were doing this in C.
\subsection{Design}
The main issue in design was the limited ability of the broker to accept and send requests to arbitrary nodes. The design of Raft calls for a centralized leader, but who this leader is is completely arbitrary and left up to chance. I was able to essentially force a node to be the leader by starting it first then making it time out very early, but if I wanted that node to fail, I'd have no idea who the next leader was, so I'd be unable to hard-code that into the broker. The solution, of course, was to write a client node to sit between the broker and the cluster to handle all the redirects, and this seems to replicate real applications in that an extra layer of abstraction between the actual user and your server can be a very useful thing.

Initially, when I was considering ideas for this project, I decided upon a Scatter-style\footnote{Glendenning et al., \href{http://homes.cs.washington.edu/~arvind/papers/scatter.pdf}{Scalable consistency in Scatter}. 2011} combination of Raft and Chord, where the DHT overlay would use the Chord topology, and each ``node'' in the DHT would be a Raft cluster of 5 nodes. This turned out to be unfeasible due to time constraints, but I did learn a few things from thinking about how to integrate the two. First, Scatter's approach to getting the Raft/Paxos clusters to agree turns out to be widely applicable -- with the Chord overlay, the only modification needed is to use 2PC to make the changes consistent with neighboring clusters, then broadcast the change to everyone in the finger table so it can propagate around the ring faster. In addition, the issue of making sure the changes maintain consistency in the face of failures is simplified by the fact that replication is taken care of within the Raft/Paxos clusters, so different DHT nodes with copies of the same key won't need to worry about using the DHT protocols to keep themselves consistent. Given what I've implemented already, this probably won't be too much additional work -- making a new subclass of \verb|RaftBaseNode| with the rather simple Chord DHT protocols built in would be sufficient.
\end{document}
